\cursedforest\ - A Random Forest Implementation for ``Big'' and ``Wide'' Data \\ Supplementary Information

### 2016-12-02 doc/publications/plos16/plos16.tex
changes to  plos16.tex

\section{Conclusion}
We have demonstrated that using a different model of paralleization can extend the random forest to delaing with the
case of an extremely large number of variables. By comparing this implementation to other implentations (including those
optimized for large data sets) we have demonstrated the utility of this implentation.

We have treated the case of variable selection in a $p >> n$ model where most of the variables are uninformative. We
have also demostrated the utility of the model for large GWAs data sets.

@@

Donoho and Tanner \cite{Donoho.and.Tanner.2009} give a ``universal phase change'' result based on arguments from
combinatorial geometry. This has applications in a large number of areas including variable selection in high
diensions. Consider figure \ref{figure:phase-diagram-equivalence.png} which shows the region where a model can recover
the important variables, plotted as a function of $\delta = n/p$ and $\rho =k/n$ (where $k$ is the number of significant
variables). There is a distinct doundary, shown empirically and also based on arguments from combinatorial geometry, to
the region where we can reliably recover significant variables.

@@

Notes:
\begin{itemize}
\item \cite{Donoho.and.Stodden.2006} evalute variable selection by 
$$\frac{\Vert\hat{\beta}-\beta\Vert_2}{\Vert\beta\Vert_2}$$
However there is no $\hat{\beta}$ in a tree model. Piotr was using vaiable importace as
a measure of $\hat{\beta}$ (i.e. we expect varible importance to be proportional to $\hat{\beta}$. Is this correct?

\end{itemize}

>>>>

\section{Conclusion}
We have demonstrated that using a different paralleization model can extend random forests to the case of an
extremely large number of variables. We have treated the case of variable selection in a $p >> n$ model where most of
the variables are uninformative and have demonstrated the utility of the model for large GWAs data sets. By comparing
this implementation to other implementations (including those optimized for large data sets) we have demonstrated the
utility of this approach.

@@

Donoho and Tanner \cite{Donoho.and.Tanner.2009} give a ``universal phase change'' result that has applications in a
large number of areas including variable selection in high dimensions. Consider figure
\ref{figure:phase-diagram-equivalence.png} which shows the region where a model can recover the important variables,
plotted as a function of $\delta = n/p$ and $\rho =k/n$ (where $k$ is the number of significant variables). There is a
distinct boundary, shown empirically and also based on arguments from combinatorial geometry, to the region where we can
reliably recover significant variables.  \cite{Donoho.and.Stodden.2006} investigate the behaviour of a number of
regression approaches for variable selection (lars, lasso and forward stepwise) and make the point that above the
phase-transition line variable recovery is still possible by a combinatorial approach.

It is not surprising that that it is more difficult to recover the signal variable in the upper-left area of the figure,
as the problem is both under-determined and sparse. What is surprising is the connection with arguments from
combinatorial geometry. This suggests that we are seeing a universal rule rather than an implementation issue. As
\cursedforest is designed for extremely large numbers of variables it is likely to be operating in difficult regions of
the figure.

@@

Notes:
\begin{itemize}
\item \cite{Donoho.and.Stodden.2006} evaluate variable selection by 
$$\frac{\Vert\hat{\beta}-\beta\Vert_2}{\Vert\beta\Vert_2}$$
However there is no $\hat{\beta}$ in a tree model. We are using variable importance as
a measure of the ordering of $\hat{\beta}$ (i.e. we expect variable importance to be proportional to $\hat{\beta}$). Is this correct?
\item as RF is sort of greedy approximation to combinatorial approach, it may have some success in the dead zone
\item  \cite{Donoho.and.Stodden.2006} are considering problems with $L_2$ errors (with additional  $L_1$ norms on the
  $\beta$. $L_1$).
How will their results carry over to 0-1 loss?
\end{itemize}
